---
title: "Pursue STEM activity"
subtitle: "Visualizing"
output: learnr::tutorial
css: "css/learnr_303.css"
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
# These are the packages you will need for this activity
packages_needed <- c("tidyverse", "learnr", "DT")

package.check <- lapply(
  packages_needed,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE, repos = "http://cran.us.r-project.org")
    }
  }
)

# Credit: package.check based on a helpful post from Vikram Baliga https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/

library(learnr)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)

track_data <- read_csv("https://raw.githubusercontent.com/elb0/pursuestem_activity/master/data/for_students.csv")

```

## Instructions

In this activity you're going to be choossing one or more of the artists you/your fellow students suggested and explore the 'mood' of their music. 

```{r, echo = FALSE}
track_data %>% 
  filter(artist_name %in% c("BTS")) %>% 
  ggplot(aes(x = valence, y = energy, colour = artist_name)) +
    geom_point() +
    theme_minimal() +
    ylim(0,1) +
    xlim(0,1) +
    geom_hline(yintercept = 0.5) +
    geom_vline(xintercept = 0.5) +
    xlab("Positiveness (valence)") +
    ylab("Engergy") + 
    labs(title = "Song moods based on energy and positiveness", 
         subtitle = "Using data from the Spotify API", 
         colour = "Artist(s)") +
  scale_color_manual(values = c("darkorchid4", "chartreuse4", "cyan2", "firebrick1"))
```


## Data dictionary

|variable                 |class     |description |
|:---|:---|:-----------|
|artist_name             |character | Song artist|
|artist_popularity       |number | Popularity score (0-100) based on track play and recency. Higher is better |
|artist_followers        |number | Number of people who 'follow' the artist on Spotify|
|album_name              |character| Name of the album the song if from |
|album_release_year      |number | Year of album release|
|track_name              |character | Song name|
|danceability            |number    | Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable. |
|energy                   |number    | Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy. |
|loudness                 |number    | The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typical range between -60 and 0 db.|
|speechiness              |number    | Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks. |
|acousticness             |number    | A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.|
|instrumentalness         |number    | Predicts whether a track contains no vocals. “Ooh” and “aah” sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly “vocal”. The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content. Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0. |
|liveness                 |number    | Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live. |
|valence                  |number    | A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry). |
|tempo                    |number    | The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration. |
|mode_name                    |character    | Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. |
|duration_ms                    |number    | Length of the song in miliseconds. |
|explicit                       |logical   | Is the song rated explicit? |
|artist_genre                   |character | This variable is still a little messy, but you can see the genres that the artist's music is considered to belong to.|



## Raw data

```{r}
DT::datatable(track_data)
```

## Your turn!

```{r table, exercise = TRUE, exercise.setup = "setup", exercise.lines = 17}
track_data %>% 
  filter(artist_name %in% c("BTS", "GOT7")) %>% 
  ggplot(aes(x = valence, y = energy, colour = artist_name)) +
    geom_point(alpha = 0.5) +
    theme_minimal() +
    ylim(0,1) +
    xlim(0,1) +
    geom_hline(yintercept = 0.5) +
    geom_vline(xintercept = 0.5) +
    xlab("Positiveness (valence)") +
    ylab("Engergy") + 
    labs(title = "Song moods based on energy and positiveness", 
         subtitle = "Using data from the Spotify API", 
         colour = "Artist(s)",
         caption = "Created by [your name] as part of Pursue STEM") 
```


## More things you could try


### Add more artists

In the second line of code, add another artist. Make sure the spelling and capitalisation is extactly as it is in the raw data!

```
  filter(artist_name %in% c("Taylor Swift", "Bruno Mars")) %>% 
```  

### Change colours
Try adding the following code:
```
# only works if you have one or three artists or fewer
scale_color_manual(values = c("darkorchid4", "chartreuse4", "cyan2", "firebrick1"))
```

Change colours, here are some names you can tell R: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

### Make separate plots
Try adding the following code:
```
facet_wrap(~arist_name)
```
